Global Climate and Atmosphere Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2018-01-04 ms.tex
amp off

@@

\subsection{Goodness of Fit}
\label{sec:fitquality}
We performed several tests of the quality of the light curve fits.  First we predicted the level of scatter in the light curves based on photon noise alone, then compared this value to the root-mean-square (rms) of the fit residuals.  The Spitzer 4.5 $\mu$m light curve rms is within 10\% of the expected photon noise limit (691 versus 640 ppm), whereas the 3.6 $\mu$m light curve has significantly larger rms (797 versus 470 ppm), due to time-correlated red noise (discussed below). The expected photon-limited rms for the WFC3 spectrosopic light curves ranges from 430 - 530 parts per million (ppm), and the measured rms was typically within 5\% of expectations for all spectroscopic channels.  For the WFC3 band- integrated white light curve, the rms was slightly larger than predicted (175 versus 122 ppm). There are a number of possible origins for this discrepancy, including imperfect background subtraction, variation in the position of the spatial scan on the detector, loss of flux outside the extraction window, and imperfect modeling of the astrophysical signal or instrument systematics. However, our analysis of the composition and thermal structure of the planet's atmosphere is based on the wavelength-dependent data (where we reach the photon limit), so we do not further explore the small excess noise in the white light curve.

@@

\subsection{Hotspot Offset}
The location of peak brightness of the phase curve is a proxy for the efficiency of thermal reradiation to space and relative to heat transport by advection (the radiative-to-advective timescale). The faster the reradiation, the less heat can be redistributed to the nightside. We compared the measured hotspot offset for the different models and found FIXME.

%\subsection{Clouds?}

\section{Atmospheric Composition and Thermal Structure}
\label{sec:composition}
To interpret the phase-resolved spectra, we performed two model fits. First, we modeled the planet flux as a simple blackbody to estimate the dayside brightness temperature and test for significant absorption or emission features. We then performed a more sophisticated grid-based retrieval to evaluate the atmospheric composition and thermal structure.

@@

\begin{deluxetable}{lll}
	\tablecolumns{3}
	\tablewidth{0pt}:
	\tablecaption{Phase-resolved Effective Temperature \label{table:teffs}}
	\tablehead{
	\colhead{Phase} & \colhead{Effective Temperature} & \colhead{$\chi^2_\nu$}}
		\startdata
		0.1 & $ 1919 \pm 40 $ & 0.6 \\
		0.2 & $ 2256 \pm 24 $ & 0.8 \\
		0.3 & $ 2651 \pm 17 $ & 1.2 \\
		0.4 & $ 2906 \pm 11 $ & 0.8 \\
		0.5 & $ 3013 \pm 8 $ & 1.6 \\
		0.6 & $ 2885 \pm 12 $ & 1.3 \\
		0.7 & $ 2636 \pm 16 $ & 0.9 \\
		0.8 & $ 2313 \pm 23 $ & 1.1 \\
		0.9 & $ 2027 \pm 36 $ & 0.5 \\
		\enddata
		\vspace{-0.8cm}
		\tablecomments{comments}
	\end{deluxetable}
%Over the WFC3 bandpass, the phase-resolved spectra are featureless and consistent with blackbodies (with best fit $\chi^2_\nu$ ranging from $0.5-1.3$). However, the \Spitzer\ data, particularly at 4.5 $\mu$m, are discrepant with the best fit blackbody by $3-5\,\sigma$ at most orbital phases, indicating the presence of infrared molecular absorption or emission features. We investigate possible atmospheric compositions and temperature structures in \S\,\ref{sec:composition}.

@@

Figure\,\ref{fig:opacities} shows the opacity contributions of key absorbers for the best fit.  In the optical (which we do not observe directly), there is strong absorption by TiO, VO, and FeH.  In the near-infrared, the H$_2$O, H-, and hydrides/oxides all contribute to the opacity, leading to nearly constant opacity over the WFC3 wavelength range.  In cooler atmospheres, water is the dominant absorber over this bandpass, but in WASP-103b, H$_2$O is partially dissociated in the photosphere, leading to a drop in abundance by a factor of $\sim10$ (see Figure\,\ref{fig:summary}). Water also has intrinsically weaker features at high temperature \citep{tinetti12}. Finally, water is an efficient coolant, so lower water abundance leads to a more isothermal photosphere, which further weakens the contrast in and out of the water band in the observed spectrum. On top of this, there is significant H- opacity from single H atoms bound with free electrons, which fills in the opacity at wavelengths shorter than $1.5\,\mu$m.  Taken together, all these factors conspire to produce a nearly featureless spectrum from $1.1 - 1.7\,\mu$m. Finally, in the infrared the dominant absorber is CO, which produces the emission feature at \Spitzer\ wavelengths.

@@

Figure\,\ref{fig:summary} shows a summary of the temperature-pressure profile and abundances for the best fit model. The T-P profile is inverted, with temperature increasing from 2800 to 3500 K over the pressure range $10^{-2} - 10^{-3}$ bar.  The  thermal inversion is driven by absorption of optical light by oxides and hydrides in the upper atmosphere.  The contribution functions FIXME
%Figure\,\ref{fig:TP} shows a selection of 100 randomly selected dayside T-P profiles. The profiles are nearly isothermal over a pressure range of $1-10^{-2}$ bars, with a median temperature near 2900 Kelvin. The temperature increases over $10^{-2} - 10^{-3}$ bars to 3500 Kelvin, and then begins to decrease again at pressures below $10^{-3}$ bar.

@@

The retrieved C/O is consistent with solar, with a $1\sigma$ confidence interval of $0.54 - 0.85$. We infer an upper limit on C/O of 0.9 at $3\,\sigma$ confidence, driven by the fact that the atmospheric chemistry is expected to change dramatically when C/O exceeds unity. For a carbon-rich composition, the equilibrium abundance of methane relative to CO increases by orders of magnitude compared to an oxygen-rich  composition \citep[e.g.][]{madhusudhan11}. Our \Spitzer\ eclipse depths are sensitive to the relative abundance of these species, so we can confidently rule out a carbon-rich composition in spite of the lack of spectrally resolved features.

@@

Figure\,\ref{fig:gcmcomparison} shows GCM predictions compared to the best fit model to the WFC3 phase curve. The WFC3 bandpass is near the peak of the planet's thermal emission, so it provides a good test of whether the GCM reproduces the properties of the planet.  Our nominal model was a cloud-free, solar composition atmosphere with TiO/VO opacity and no added drag. This model qualitatively reproduces some features of the data, such as the large day-night temperature contrast; however, it predicts a larger offset in peak brightness from the substellar point than we observe.

To achieve a better fit to the data, we varied the nominal model parameters. We tested models with enhanced metallicity ([Fe/H] = 0.5), no TiO/VO, and added atmospheric drag.  Lorentz drag can occur in the hottest exoplanet atmospheres, which may be partially ionized and coupled to the background magnetic field. We parameterized the drag (\textbf{Vivien: how did you model drag?}), with timescales were $\tau_\mathrm{drag} = 10^3, 10^4$ and $10^6$ s.

The models with added drag more closely reproduce the small hotspot offset that is observed. The main effect of drag is to slow down advection of heat to the nightside, which shifts the peak brightness closer to the substellar point.  The best fit WFC3 phase curve has a hotspot offset of FIXME. The drag models have offsets of $0.2$ and $2.3$ degrees for $\tau_\mathrm{drag} = 10^3$ and $10^4$ s, compared to $15$ degrees for the nominal GCM.  The drag models also provide a good match to the observed dayside flux, for drag timescales in the range $\tau_\mathrm{drag} = 10^3 - 10^4$ s.  However, the drag models underestimate the flux at longitudes away from the substellar point, predicting 20\% less flux than is observed at quadrature. \textbf{Vivien: are these fast drag timescales even reasonable?}).

@@

>>>>

\subsection{Best Fits and Uncertainties}
To determine the best fits, we performed a least-squares $\chi^2$ minimization for each wavelength and model. For a subset of these cases, we also performed a Markov chain Monte Carlo (MCMC) analysis to estimate parameter uncertainties. We used \texttt{emcee} \citep{foremanmackey13} to fit the \HST/WFC3 light curves and differential evolution Monte Carlo for the \Spitzer\ fits \citep{braak06}. We ran the MCMC until convergence according to the Gelman-Rubin statistic FIXME. MCMC techniques only produce robust uncertainties when the noise is normally distributed and white, so to account for correlated noise in the $3.6\,\mu$m light curve we fit the wavelet model from \cite{carter09} (simultaneously with the other model parameters). We used a Haar wavelet and let the power spectral density of the red noise vary, following \cite{diamond-lowe14}. In our final fit, the noise power spectrum $1/f^\gamma$ had $\gamma = 1.1 \pm 0.1$.

@@

\subsection{Goodness of Fit}
\label{sec:fitquality}
We performed several tests of the quality of the light curve fits.  First we predicted the level of scatter in the light curves based on photon noise alone, then compared this value to the root-mean-square (rms) of the fit residuals.  The Spitzer 4.5 $\mu$m light curve rms is within 10\% of the expected photon noise limit (691 versus 640 ppm), whereas the 3.6 $\mu$m light curve has significantly larger rms (797 versus 470 ppm), due to time-correlated red noise (discussed below). The expected photon-limited rms for the WFC3 spectrosopic light curves ranges from 430 - 530 parts per million (ppm), and the measured rms was typically within 5\% of expectations for all spectroscopic channels.  For the WFC3 band-integrated white light curve, the rms was slightly larger than predicted (175 versus 122 ppm). There are a number of possible origins for this discrepancy, including imperfect background subtraction, variation in the position of the spatial scan on the detector, loss of flux outside the extraction window, and imperfect modeling of the astrophysical signal or instrument systematics. However, our analysis of the composition and thermal structure of the planet's atmosphere is based on the wavelength-dependent data (where we reach the photon limit), so we do not further explore the small excess noise in the white light curve.

@@

\subsection{Phase Curve Amplitudes and Hotspot Offsets}
The shape of a phase curve can be summarized with two parameters: the amplitude of thermal phase variation (minimum to maximum brightness, divided by the secondary eclipse depth) and the location of peak brightness (typically called a ``hotspot offset" and measured in degrees eastward of the substellar point). Larger amplitudes and smaller offsets are both indicative of inefficient heat transport to the nightside.

Table\,\ref{table:amps_offsets} lists the estimated amplitudes and offsets for the phase curve fits (median and $1\,\sigma$ credible interval). These values are from the sine curve model for the thermal phase variation, which directly fits for the amplitude and hotspot offset. For the WFC3 phase curves, the amplitudes are consistent with unity and the offsets are consistent with zero. We note that some of the spectroscopic channels have amplitudes greater than unity, which is physically impossible (the planet cannot emit negative flux). As a sanity check, we take the weighted average over the WFC3 channels and find an average amplitude of $0.91\pm0.02$, which shows that at higher precision the results are physical and agree well with the large amplitudes observed for other hot planets \citep{komacek16}.  The \Spitzer\ phase curves also have large amplitudes (near 0.85) and hotspot offsets consistent with zero.

\begin{deluxetable}{LLL}
	\tablecolumns{3}
	\tablewidth{0pt}:
	\tablecaption{Phase Curve Properties \label{table:amps_offsets}}
	\tablehead{
	\colhead{Wavelength} & \colhead{Amplitude} & \colhead{Offset (degrees)}}
		\startdata
		1.175 & 1.11 \pm 0.27 & 1.8 \pm 1.3 \\
		1.225 & 0.88 \pm 0.17 & -1.1 \pm 1.4 \\
		1.275 & 0.92 \pm 0.17 & -0.6 \pm 1.4 \\
		1.325 & 0.82 \pm 0.14 & -1.8 \pm 1.5 \\
		1.375 & 0.91 \pm 0.16 & -1.6 \pm 1.4 \\
		1.425 & 0.83 \pm 0.13 & -2.5 \pm 1.4 \\
		1.475 & 1.04 \pm 0.19 & -0.3 \pm 1.2 \\
		1.525 & 0.99 \pm 0.19 & 0.3 \pm 1.3 \\
		1.575 & 0.86 \pm 0.18 & 1.8 \pm 1.2 \\
		1.625 & 0.89 \pm 0.17 & 0.3 \pm 1.3 \\
		3.6 & 0.86 \pm 0.13 & 2.0 \pm 0.7 \\
		4.5 & 0.83 \pm 0.05 & 1.0 \pm 0.4 \\
		\enddata
\end{deluxetable}

%\subsection{Clouds?}
\section{Atmospheric Composition and Thermal Structure}
\label{sec:composition}
To interpret the phase-resolved spectra, we performed two model fits. First, we modeled the planet flux as a simple blackbody to estimate the dayside brightness temperature and test for significant absorption or emission features. We then performed a more sophisticated grid-based retrieval to evaluate the atmospheric composition and thermal structure.

@@

\begin{deluxetable}{LLL}
	\tablecolumns{3}
	\tablewidth{0pt}:
	\tablecaption{Phase-resolved Effective Temperature \label{table:teffs}}
	\tablehead{
	\colhead{Orbital Phase} & \colhead{$T_\mathrm{eff}$} & \colhead{$\chi^2_\nu$}}
		\startdata
		0.06 - 0.15 & 1919 \pm 40 & 0.6 \\
		0.15 - 0.25 & 2256 \pm 24 & 0.8 \\
		0.25 - 0.35 & 2651 \pm 17 & 1.2 \\
		0.35 - 0.44 & 2906 \pm 11 & 0.8 \\
		0.44 - 0.56 & 3013 \pm 8 & 1.6 \\
		0.56 - 0.65 & 2885 \pm 12 & 1.3 \\
		0.65 - 0.75 & 2636 \pm 16 & 0.9 \\
		0.75 - 0.85 & 2313 \pm 23 & 1.1 \\
		0.85 - 0.94 & 2027 \pm 36 & 0.5 \\
		\enddata
		\vspace{-0.8cm}
		\tablecomments{comments}
	\end{deluxetable}

@@

Figure\,\ref{fig:opacities} shows the opacity contributions of key absorbers for the best fit model.  In the optical (which we do not observe directly), there is strong absorption by TiO, VO, and FeH.  In the near-infrared, H$_2$O, H-, and hydrides/oxides all contribute to the opacity, leading to nearly constant opacity over the WFC3 wavelength range.  In cooler atmospheres, water is the dominant absorber over this bandpass \citep[e.g.][]{kreidberg14b, line16}, but in WASP-103b, H$_2$O is partially dissociated in the photosphere, leading to a drop in abundance by a factor of $\sim10$ (see Figure\,\ref{fig:summary}). Water also has intrinsically weaker features at high temperature \citep{tinetti12}. Water is also an efficient coolant, so lower water abundance leads to a more isothermal photosphere, which further weakens the contrast in and out of the water band. On top of this, there is significant H- opacity from single H atoms bound with free electrons, which fills in the opacity at wavelengths shorter than $1.5\,\mu$m.  Taken together, all these factors conspire to produce a nearly featureless spectrum from $1.1 - 1.7\,\mu$m. Finally, in the infrared the dominant absorber is CO, which produces the emission feature at \Spitzer\ wavelengths.

@@

Figure\,\ref{fig:summary} shows a summary of the temperature-pressure profile and abundances for the best fit model. The T-P profile is inverted, with temperature increasing from 2800 to 3500 K over the pressure range $10^{-2} - 10^{-3}$ bar.  The  thermal inversion is driven by absorption of optical light by oxides and hydrides in the upper atmosphere.  The observations are sensitive to pressures of $\sim0.01 - 0.001$ bar, which spans the tropopause, where temperature begins to increase and the water abundance drops by more than an order of magnitude.

@@

The retrieved C/O is consistent with solar, with a $1\sigma$ confidence interval of $0.54 - 0.85$. We infer an upper limit on C/O of 0.9 at $3\,\sigma$ confidence, driven by the fact that the atmospheric chemistry is expected to change dramatically when C/O exceeds unity. For a carbon-rich composition, the equilibrium abundance of methane relative to CO increases by orders of magnitude compared to an oxygen-rich  composition \citep[e.g.][]{madhusudhan11}. Our \Spitzer\ eclipse depths are sensitive to the relative abundance of these species, so we can confidently rule out a carbon-rich composition despite the lack of spectrally resolved features.

@@

Figure\,\ref{fig:gcmcomparison} shows GCM predictions compared to the best fit model to the WFC3 phase curve. The WFC3 bandpass is near the peak of the planet's thermal emission, so it provides a good test of whether the GCM reproduces the properties of the planet.  Our nominal GCM was a cloud-free, solar composition atmosphere with TiO/VO opacity and no added drag.  We also ran models with enhanced metallicity ([Fe/H] = 0.5), no TiO/VO, and added atmospheric drag.  Lorentz drag can occur in the hottest exoplanet atmospheres, which may be partially ionized and coupled to the background magnetic field. We parameterized the drag (\textbf{Vivien: how did you model drag?}), with timescales were $\tau_\mathrm{drag} = 10^3, 10^4$ and $10^6$ s.

This model qualitatively reproduces some features of the data, such as the large day-night temperature contrast; however, it predicts a larger offset in peak brightness from the substellar point than we observe. 
The models with added drag more closely reproduce the small hotspot offset that is observed. The main effect of drag is to slow down advection of heat to the nightside, which shifts the peak brightness closer to the substellar point.  The best fit WFC3 phase curve has a hotspot offset of FIXME. The drag models have offsets of $0.2$ and $2.3$ degrees for $\tau_\mathrm{drag} = 10^3$ and $10^4$ s, compared to $15$ degrees for the nominal GCM.  The drag models also provide a good match to the observed dayside flux, for drag timescales in the range $\tau_\mathrm{drag} = 10^3 - 10^4$ s.  However, the drag models underestimate the flux at longitudes away from the substellar point, predicting 20\% less flux than is observed at quadrature. \textbf{Vivien: are these fast drag timescales even reasonable?}).

@@

what do the hot spot offsets and amplitudes mean?
